---
title: "MathVisualization"
output: 
  flexdashboard::flex_dashboard:
    vertical-layout: scroll
runtime: shiny
---




```{r "setup"}
options(shiny.sanitize.errors=FALSE)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)


library(tidyverse)
library(qualtRics)
library(corrplot)
library(ggwordcloud)
library(forcats)


camel <- function(x){ #function for camel case
    capit <- function(x) paste0(toupper(substring(x, 1, 1)), substring(x, 2, nchar(x)))
    sapply(strsplit(x, "\\."), function(x) paste(capit(x), collapse=""))
}


longify <- function(d){
  pivot_longer(d, 1:ncol(d))
}

freq_table <- function(d, prompt = "prompt"){
  (d  %>% group_by_at(1) 
  %>% summarize(ct= n()) 
  %>% mutate(prop = ct/sum(ct)
             , prompt = prompt
             )
  %>% arrange(-ct)
  %>% na.omit()
  )
}


#math_survey <- read_survey("~/Dropbox/Math Hallucinations/MathVisFullData20210802_anonymized.csv")
math_survey <- read_survey("~/Downloads/Mathematical Visualization_May 22, 2024_17.34.csv")
math_survey <- math_survey %>% filter(math_ed %in% c("Masters Degree in Math", "Ph.D. in Math", "Some Graduate Coursework"))
#math_survey <- math_survey %>% filter(Progress == 100)
math_survey <- math_survey %>% filter(Progress >= 98)
yesses <- function(v, yes="Yes"){
  mean(v==yes, na.rm=T)
}
math_survey <- math_survey %>% filter(domains != "jhgfds")
math_survey <- math_survey %>% mutate(platonist = platonist_1) %>% select(-platonist_1)

#n_bad_rows = 17

#math_survey <- math_survey %>% slice(-c(1:n_bad_rows))
n_bad_rows = 0
```


```{r "Data Cleaning from Qualtrics", include=FALSE}  

# Make long

# Remove partial data

n_subjects <- length(math_survey$Progress)

```

# Notes

### Notes

Don't forget to include "symbolicvis", which is currently commented out

Also uncomment out the ggplots, once the plots are debugged.

Uncomment "describe"

Uncomment the "noise and business" sections
# Demographics

## Demographics 

### Education

```{r "Demographics"}
education_level <- math_survey %>% group_by(math_ed) %>% summarize(ct= length(math_ed)) %>% mutate(math_degree = math_ed %>% gsub("in Math", "", .) %>% gsub("Some Graduate Coursework", "Graduate\nClasses", .) %>% gsub("Masters Degree", "Masters", .))
#renderTable({education_level})
renderPlot({
  (ggplot(education_level, aes(x=math_degree, y=ct, fill=math_degree)) + geom_col() + theme_bw(base_size=15) 
   #+ coord_flip()
   + xlab("Math Experience") + ylab("Number of Participants") + theme(legend.position = "none")+scale_fill_viridis_d()
   )
  })

```

### Domains 
```{r "Demographic Domains"}
domains <- math_survey %>%rowwise %>% mutate(domain = ((strsplit(domains, ",")))) %>% unnest(domain) %>% mutate(domain = tolower(trimws(domain)))
domains$domain <- (domains$domain %>% gsub("number tgeory", "number theory", .) %>% gsub("but: ", "", .) %>% gsub("/geometric group theory", "", .) %>% gsub("as a math education phd i am a generalist", "math education" ,. ) %>% gsub("k–8 education", "math education", .) %>% gsub(" and pde", "", .) %>% gsub("higher category", "category", .) %>% gsub("surreal numbers general relativiy etc", "surreal numbers", .) %>% gsub("/ model theory", "", .))
domains <- domains %>% group_by(domain) %>% summarize(ct=length(domain)) %>% arrange(-ct)
#renderTable({domains})
renderPlot({
ggplot(domains, aes(label = domain, size = ct^1.5, color = ct)) +
    geom_text_wordcloud_area() +
    scale_size_continuous(range = c(5, 10)) +  # Adjust the range for desired size
    theme_minimal() +
    labs(title = "Word Cloud of Domains") +
    theme(plot.title = element_text(hjust = 0.5))
})
```

### Self-reported Platonism
```{r}

#renderTable({math_survey %>% group_by(platonist) %>% summarize(ct= length(platonist))})

renderPlot({
  ggplot(math_survey, aes(x = platonist/100)) +
  stat_density(geom="line", color = "blue", size = 1.2) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "lightblue", alpha = 0.7) +
  scale_y_continuous(name = "Histogram") +
  scale_x_continuous(limits=c(0, 1), labels=scales::percent)+
  theme_bw(base_size=20) +
  labs(title = "",
       x = "Proportion of Time") +
  theme(plot.title = element_text(hjust = 0.5))

  
})

```



## Drawing

### Drawing Exactly Proportions

```{r 'Exact drawing proportions'}

drawing_proportions <- math_survey %>% summarize(
  draw_prop = yesses(draw)
)

renderTable({longify(drawing_proportions)})


```

### Draw For Whom?

```{r 'drawing for'}


renderTable({
  freq_table(math_survey 
   %>% select(draw_for)
   , "draw_for"
) %>% select(-prompt)
  })


```


### Visualization Presence

```{r "Estimate proportion of non-visualizers (pre-registered), then separate them" } 


visualization_summary <- math_survey %>% summarize(
  proportion_zero = mean(prop_of_time_1==0, na.rm=T)
  , average_time = mean(prop_of_time_1[prop_of_time_1>0], na.rm=T)
  , sd = sd(prop_of_time_1[prop_of_time_1>0], na.rm=T)
)

renderTable({longify(visualization_summary)})

```



# Sensations

## Column 
### Sensation Distribution

```{r}
sensory_characteristics <- math_survey %>% summarize(
  color = mean(visualspatial_1, na.rm=T)
  , shape = mean(visualspatial_2, na.rm=T)
  , relative_location  = mean(visualspatial_3, na.rm=T)
  , tactile = mean(sensory_1, na.rm=T)
  , auditory = mean(sensory_2, na.rm=T)
  , olfactory = mean(sensory_3, na.rm=T)
#  , touchable = 100*yesses(touch)
#  , symbols = 100*yesses(symbols)
  , dynamic = mean(dynamic_1, na.rm=T)
  #, just_symbols = mean(symbolicvis=="Yes", na.rm=T)
)



#renderTable({longify(sensory_characteristics)})

```

```{r}

sensory_characteristics_table <- (
  math_survey %>% mutate(
    color = visualspatial_1
    , shape = visualspatial_2
    , location  = visualspatial_3
    , tactility = sensory_1
    , sound = sensory_2
    , olfaction = sensory_3
    , movement = dynamic_1
  
  ) 
  %>% select(ResponseId, color:movement)
  %>% pivot_longer(cols=c(color:movement), names_to="sensory_channel", values_to="value")
  %>% group_by(sensory_channel)
  %>% summarize(mean_yes = mean(value>0, na.rm=T)
                , mean_val = mean(value[value>0], na.rm=T))
  %>% arrange(-mean_yes)
)
renderPlot({
(
  ggplot(sensory_characteristics_table %>% arrange(-mean_yes)
         , aes(x=fct_reorder(sensory_channel, mean_yes), y=mean_yes, fill=sensory_channel, label=sensory_channel
               , width=mean_val/100))
  + geom_col(aes(size=mean_val))
  + coord_flip()
  + theme_bw(base_size = 16)
  + theme(legend.position="none")
  + scale_x_discrete("Experienced Property")
  + scale_y_continuous("% reporting sensation", labels=scales::percent, limits=c(0,1))
  # + geom_text(nudge_x=-0.05, nudge_y = -0.05)
)     
})
```

### draw time proportion


```{r}


renderPlot({
  ggplot(math_survey, aes(x = prop_of_time_1/100)) +
  stat_density(geom="line", color = "blue", size = 1.2) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "lightblue", alpha = 0.7) +
  scale_y_continuous(name = "Histogram") +
  scale_x_continuous(limits=c(0, 1), labels=scales::percent)+
  theme_bw(base_size=20) +
  labs(title = "",
       x = "Proportion of Time") +
  theme(plot.title = element_text(hjust = 0.5))

  
})

```



## Touch Yes and No

### Touch Yes

```{r}

renderTable({math_survey %>% group_by(touch_1_TEXT) %>% summarize(ct= length(touch_1_TEXT))})

```

### Touch No 
```{r}


renderTable({math_survey %>% group_by(touch_2_TEXT) %>% summarize(ct= length(touch_2_TEXT))})

```

## Control

### Symbols Yes

```{r}

renderTable({math_survey %>% group_by(symbols_1_TEXT) %>% summarize(ct= length(symbols_1_TEXT))})

```

### Symbols No 
```{r}


renderTable({math_survey %>% group_by(symbols_2_TEXT) %>% summarize(ct= length(symbols_2_TEXT))})

```

# Content

## Content 0

### Proportions
```{r 'content patterns'}

content_patterns <- math_survey %>% summarize(
   prop_constraint = yesses(constraint)
  

)

renderTable({longify(content_patterns)})


```
## Content 1


### Constraints

```{r}

renderTable({bind_rows(freq_table(math_survey %>% select(constraint_1_TEXT), "Constraint Yes")
  , freq_table(math_survey %>% select(constraint_2_TEXT), "Constraint No"))})

```



### Challenges

```{r}
challenges <- math_survey %>%rowwise %>%  mutate(challenge = ((strsplit(challenges, ",")))) %>% unnest(challenge) %>% mutate(challenge = trimws(challenge))


renderTable({
  (challenges %>% select(challenge) %>% freq_table() %>% mutate(prop=ct/n_subjects))
  })

#challenges_indiv_ct <- challenges %>% group_by(StartDate, EndDate) %>% summarize(ct=n()) %>% ungroup()
```

### plot

```{r challenge plot}

#ggplot(challenges_indiv_ct, aes(x=ct)) + geom_histogram()

```

## Platonism

### Real

```{r}

platonism_patterns <- math_survey %>% summarize(
  object_real = mean(object_real_4, na.rm=T)
  , object_exist = mean(Q50_4, na.rm=T)
  , build_or_see = mean(tail(build_see_1, -n_bad_rows), na.rm=T)
#  , build_or_see = mean(build_see_1, na.rm=T)

)

renderTable({longify(platonism_patterns)})
 # ggplot(math_survey, aes(x = build_see_1/100)) +
 #  stat_density(geom="line", color = "blue", size = 1.2) +
 #  geom_histogram(aes(y = ..density..), color = "black", fill = "lightblue", alpha = 0.7) +
 #  scale_y_continuous(name = "Histogram") +
 #  scale_x_continuous(limits=c(0, 1), labels=scales::percent)+
 #  theme_bw(base_size=20) +
 #  labs(title = "",
 #       x = "Proportion of Time") +
 #  theme(plot.title = element_text(hjust = 0.5))

```

###

```{r}

renderTable({
  freq_table(math_survey 
   %>% select(platonist)
) %>% select(-prompt)
  })


```

# Control

## Control Stats

### Dynamics Sensation

```{r 'dynamic sensation'}

control_patterns <- math_survey %>% summarize(
    zoom = mean(tail(zoom_2, -n_bad_rows), na.rm=T)
  , do_touch = 100*yesses(do_touch)
  , resist_mvmt = resist_mvmt_3 #yesses(resist_mvmt_3)
  , prop_causal =  yesses(causal)


)

renderTable({longify(control_patterns)})


```

### Visual Persistence

```{r 'persistence patterns'}

persistence_patterns <- math_survey %>% summarize(
  appear_unwanted_prop = mean(appear=="Yes", na.rm=T)
  , fade_unwanted_prop = mean(fade=="Yes", na.rm=T)
  , remain_present_prop = mean(stay_present=="Yes", na.rm=T)
  , cannot_stop_seeing_prop =  mean(stop_seeing=="Yes", na.rm=T)
  , annoyed =  mean(annoyed=="Yes", na.rm=T)
  , control = mean(control_1, na.rm=T)
  

)

renderTable({longify(persistence_patterns)})


```



## Explanations

### Causal Control Yes

```{r}

renderTable({math_survey %>% group_by(causal_1_TEXT) %>% summarize(ct= length(causal_1_TEXT))})

```

### Causal Control No 
```{r}


renderTable({math_survey %>% group_by(causal_2_TEXT) %>% summarize(ct= length(touch_2_TEXT))})

```



### Do Touch Yes

```{r}

renderTable({math_survey %>% group_by(do_touch_1_TEXT) %>% summarize(ct= length(do_touch_1_TEXT))})

```

### Do Touch No 
```{r}


renderTable({math_survey %>% group_by(do_touch_2_TEXT) %>% summarize(ct= length(do_touch_1_TEXT))})

```

## New column

### Uncontrollable Persistence

```{r}

renderTable({
  bind_rows(
    freq_table(math_survey 
       %>% select(appear_1_TEXT)
      , "draw_for"
      )
    ,  freq_table(math_survey 
       %>% select(fade_1_TEXT)
      , "Fade"
      )
    ,  freq_table(math_survey 
       %>% select(stay_present_1_TEXT)
      , "Stay Present"
      )
    ,  freq_table(math_survey 
       %>% select(stop_seeing_1_TEXT)
      , "Stop Seeing"
      )      
  )  %>% na.omit()
  })

```



### Cannot Stop Seeing Yes 
```{r}


renderTable({math_survey %>% group_by(stop_seeing_1_TEXT) %>% summarize(ct= length(stop_seeing_1_TEXT)) %>% arrange(-ct)})

```




### Annoyed Yes 
```{r}


renderTable({math_survey %>% group_by(annoyed_1_TEXT) %>% summarize(ct= length(annoyed_1_TEXT)) %>% arrange(-ct)})

```


# Cult. of Viv 


### Visualization Vividness



```{r "Change in vividness over time"}
# Come back here

vividness_summary <- math_survey %>% summarize(
    session_increasing_vividness = yesses(session_vivid, "More vivid")
  , session_decreasing_vividness = yesses(session_vivid, "Less vivid")
  , session_consistent_vividness = mean(session_vivid ==  "They stay consistent in their  vividness", na.rm=T)
  , session_variable_vividness = yesses(session_vivid,  "It depends on where I am in a reasoning process")+yesses(session_vivid,  "It just depends" )
  ,  domain_increasing_vividness = mean(domain_vivid == "More  vivid", na.rm=T)
  ,  domain_decreasing_vividness = mean(domain_vivid == "Less  vivid", na.rm=T)
    , domain_consistent_vividness = mean(domain_vivid ==  "They stay consistent in their vividness", na.rm=T)
  , domain_variable_vividness = mean(domain_vivid ==  "It just depends", na.rm=T)
)
renderTable({longify(vividness_summary)})
```

### Isolation Distribution

```{r}
isolation_characteristics <- math_survey %>% summarize(
  noise_noisy = mean(tail(noise_noisy_1, -n_bad_rows), na.rm=T)
  , people_company = mean(tail(people_company_1, -n_bad_rows), na.rm=T)
  , hands_active = mean(tail(hands_active_1, -n_bad_rows), na.rm=T)
  , light_bright = mean(tail(light_bright_1, -n_bad_rows), na.rm=T)
  , space_full = mean(tail(space_full_1, -n_bad_rows), na.rm=T)
  , surface_present  = mean(tail(surface_need_1, -n_bad_rows), na.rm=T) # one-sided

)

renderTable({longify(isolation_characteristics)})


```


```{r}
isolation_characteristics_by_subjects <- (
  math_survey
  %>% mutate(busyness_need = noise_noisy_1
                                +people_company_1
                                +hands_active_1
                                +space_full_1
                                +light_bright_1
             , pickiness = (abs(noise_noisy_1 - 50)
                            +abs(people_company_1 - 50)
                            +abs(hands_active_1 - 50)
                            +abs(space_full_1 - 50)
                            +abs(light_bright_1 - 50)
                            + abs(surface_need_1/2)

             )
  )
  %>% select(busyness_need, pickiness)
)

renderTable(isolation_characteristics_by_subjects)

```


# Describe 

## Column

### Descriptions of Visualizations

```{r Descriptions}
# freq_table(math_survey 
#        %>% select(describe)
#       , "describe"
#       )

```


## Vis

### Visualization specialness

```{r Vis special}
# freq_table(math_survey
#        %>% select(vis_domain
#       , "Vis Often"
#       )
# )

# renderTable(renderTable({
#   bind_rows(
#     freq_table(math_survey 
#        %>% select(vis_often_3_TEXT)
#       , "Math Special"
#       )
#     ,  freq_table(math_survey 
#        %>% select(vis_often_4_TEXT)
#       , "All Domains"
#       ))
# })
```





```{r}
math_survey_rename <- math_survey %>% rename(
  color = visualspatial_1
  , shape = visualspatial_2
  , relative_location = visualspatial_3
  , tactile = sensory_1
  , auditory = sensory_2
  , olfactory = sensory_3
  , dynamic = dynamic_1
  , object_real = object_real_4
  , object_exist = Q50_4
  , build_or_see = build_see_1
  , zoom = zoom_2
  , resist_mvmt = resist_mvmt_3
  , noise_noisy = noise_noisy_1
  , hands_active = hands_active_1
  , light_bright = light_bright_1
  , hands_active = hands_active_1
  , space_full = space_full_1
  , visualization_proportion = prop_of_time_1
  , people_company = people_company_1
  , control = control_1
  , surface_present = surface_need_1 

 
) %>% mutate( symbols = unlist(symbols)=="Yes"
              , prop_causal = unlist(causal) == "Yes"
              , domain_increasing_vividness = (domain_vivid == "More  vivid")
              , session_increasing_vividness = (session_vivid == "More vivid")
         #     , platonist = platonist %in% c("Sometimes", "Yes, always")
              , constrained = constraint %in% c("Yes")
               , annoyed = annoyed %in% c("Yes")
              , appear_unwanted = appear == "Yes"
              , fade_unwanted = fade == "Yes"
              , cannot_stop_seeing = stop_seeing == "Yes"
              , remain_present = stay_present == "Yes"
              )

math_survey_rename$build_or_see[1:n_bad_rows] <- 0/0
math_survey_rename$zoom[1:n_bad_rows] <- 0/0
math_survey_rename$noise_noisy[1:n_bad_rows] <- 0/0
math_survey_rename$people_company[1:n_bad_rows] <- 0/0
math_survey_rename$hands_active[1:n_bad_rows] <- 0/0
math_survey_rename$light_bright[1:n_bad_rows] <- 0/0
math_survey_rename$space_full[1:n_bad_rows] <- 0/0
math_survey_rename$surface_present[1:n_bad_rows] <- 0/0
 
 

msr <-  (math_survey_rename
         %>% filter(math_ed %in% c("Bachelor Degree in Math"
                                           , "Ph.D. in Math", "Masters Degree in Math"
                                           , "Some Graduate Coursework"))
         %>% select(color, shape, relative_location, tactile, auditory, olfactory
                    , symbols, dynamic, object_real, object_exist, build_or_see
                    , zoom, symbols, resist_mvmt, prop_causal, domain_increasing_vividness
                    , noise_noisy, hands_active, light_bright , hands_active, platonist
                    , visualization_proportion, space_full, annoyed, constrained
                    , session_increasing_vividness
                                       , remain_present
                                       , cannot_stop_seeing
                                       , fade_unwanted
                                       , appear_unwanted
                                       , control
                                       , people_company
                                       , surface_present)  ) 





corrplot(cor(msr, use="pairwise.complete.obs"), order='hclust')


```



```{r}
require(FactoMineR)
library(missMDA)
msr.comp <- imputePCA(msr)
msr.pca <- PCA(msr.comp$completeObs)

```